---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MargLabGHG

MargLabGHG provides a basic workflow processing data from Gasmet portable gas analyzers, including support for flux modeling and calculations.

## Installation

You can install the development version of MargLabGHG from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("redseasoils/MargLabGHG")
```

## Quick Start

To get started, you'll need:

  1. A TXT file from a Gasmet portable gas analyzer
  2. A spreadsheet (.xlsx) containing variable chamber dimensions and variables, including chamber temperature (В°C), collar height (cm; if applicable), chamber height (cm), and input and output tube lengths (cm).
  
We recommend storing these files in a specific structure so that the package's functions can find many of these files easily and process them in batches. More on that below, but for now, let's focus on getting 

```{r import-gasmet-file}
library(MargLabGHG)
# Use an example TXT file directory
txt_file_dir <- system.file("data", "00_raw", "gas_concentration", "Site1", 
                            "20240830", package = "MargLabGHG")
# And a TXT file within that directory
txt_file <- "301 BT.TXT"
# Import
gasmet_data <- import_gasmet_data(txt_file_dir, txt_file)
head(gasmet_data)

# Use an example chamber variables spreadsheet
chamber_dir <- system.file("data", "00_raw", "chamber_volume", "Site1", 
                           package = "MargLabGHG")
chamber_file <- "20240830.xlsx"
chamber <- import_chamber_volume(chamber_dir, chamber_file)
```

## File Storage for Batch Processing

Files must be structured in a standard format to batch process Gasmet data. File structure must meet the following conditions:

  1. All Gasmet TXT files must be stored in a directory in which there are no other TXT files. The files *can* be stored in subdirectories, but there must be one main directory below which the only TXT files are Gasmet files.
  2. Similarly, all spreadsheets containing chamber variables must be stored in a directory which contains no other .xlsx files (this can be the same as the TXT file directory). These files may also be stored in subdirectories. 
  
Below is an example file structure:
```bash
в””в”Ђв”Ђ Data
В В  в””в”Ђв”Ђ GHG
В В   В В  в””в”Ђв”Ђ Gasmet
В В   В В   В В  в””в”Ђв”Ђ raw
В В   В В   В В  в”‚В В  в”њв”Ђв”Ђ Site_A
В В   В В   В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 20240701
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1101.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1102.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1201.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1202.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в””в”Ђв”Ђ Chamber.xlsx
В В   В В   В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 20240708
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1101.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1102.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1201.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1202.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в””в”Ђв”Ђ Chamber.xlsx
В В   В В   В В  в”‚В В  в””в”Ђв”Ђ Site_B
В В   В В   В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 20240702
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2101.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2102.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2201.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2202.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в””в”Ђв”Ђ Chamber.xlsx
В В   В В   В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 20240709
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2101.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2102.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2201.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2202.TXT
В В   В В   В В  в”‚В В  в”‚В В  в”‚В В  в””в”Ђв”Ђ Chamber.xlsx
```

With the file structure above, assuming the working directory is the root directory containing the "Data" folder, the import functions would be:

``` {r eval=FALSE}
# Set option for TXT file directory, then import
options("gaseous.gasemt_TXT_dir" = "Data/GHG/Gasmet/raw")
import_gasmet_data()

# OR specify the path in the import function, without option setting
import_gasmet_data("Data/GHG/Gasmet/raw")

# Chamber measurements
import_chamber_volume("Data/GHG/Gasmet/raw")
```

