---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gasFluxR

gasFluxR provides a basic workflow processing data from Gasmet portable gas analyzers, including support for flux modeling and calculations.

## Installation

You can install the development version of gasFluxR from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("redseasoils/gasFluxR")
```

## Getting Started

To calculate gas flux from Gasmet data using the package's functions, you'll need:

  1. A TXT file from a Gasmet portable gas analyzer
  2. A spreadsheet (.xlsx) containing variable chamber dimensions and variables, including chamber temperature (В°C), collar height (cm; if applicable), chamber height (cm), and input and output tube lengths (cm).
  
We recommend storing these files in a specific structure so that the package's functions can find many of these files easily and process them in batches. More on that below, but for now, let's focus on getting 

```{r import-gasmet-file}
library(gasFluxR)
# Use an example TXT file directory
txt_file_dir <- system.file("data", "00_raw", "gas_concentration", "Site1", 
                            "20240830", package = "gasFluxR")
# And a TXT file within that directory
txt_file <- "301 BT.TXT"
# Import
gasmet_data <- import_gasmet_data(txt_file_dir, txt_file)
head(gasmet_data)

# Use an example chamber variables spreadsheet
chamber_dir <- system.file("data", "00_raw", "chamber_volume", "Site1", 
                           package = "gasFluxR")
chamber_file <- "20240830.xlsx"
chamber <- import_chamber_volume(chamber_dir, chamber_file)
```

## File Storage for Batch Processing

Files must be structured in a standard format to batch process Gasmet data. File structure must meet the following conditions:

  1. All Gasmet TXT files must be stored in a directory in which there are no other TXT files. The files *can* be stored in subdirectories, but there must be one main directory below which the only TXT files are Gasmet files.
  2. Similarly, all spreadsheets containing chamber variables must be stored in a directory which contains no other .xlsx files (this can be the same as the TXT file directory). These files may also be stored in subdirectories. 
  
Below is an example file structure, in which "Gasmet_Data/raw" is the main directory for both Gasmet TXT files and spreadsheets:
```bash
в””в”Ђв”ЂGasmet_Data
в”‚В В  в””в”Ђв”Ђ raw
в”‚В В  в”‚В В  в”њв”Ђв”Ђ Site_A
в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 20240701
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1101.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1102.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1201.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1202.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в””в”Ђв”Ђ Chamber.xlsx
в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 20240708
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1101.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1102.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1201.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1202.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в””в”Ђв”Ђ Chamber.xlsx
в”‚В В  в”‚В В  в””в”Ђв”Ђ Site_B
в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 20240702
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2101.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2102.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2201.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2202.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в””в”Ђв”Ђ Chamber.xlsx
в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 20240709
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2101.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2102.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2201.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2202.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в””в”Ђв”Ђ Chamber.xlsx
```

With the file structure above, assuming the working directory is the root directory containing the "Data" folder, the import functions would be:

``` {r eval=FALSE}
# Set option for TXT file directory, then import
options("gaseous.gasemt_TXT_dir" = "Data/GHG/Gasmet/raw")
import_gasmet_data()

# OR specify the path in the import function, without option setting
import_gasmet_data("Data/GHG/Gasmet/raw")

# Chamber measurements
import_chamber_volume("Data/GHG/Gasmet/raw")
```


## Flux Modeling Procedures

The function `model_gas_flux()` detects the gas species via the column name in the `ppm_var` argument (or using `gas_name` directly, if specified). The data are dispatched to modeling procedures based on species:

  + **CO~2~**: Linear model with four options for deadband (i.e. initial chamber
  disturbance) removal:
    1. None: All data are modeled
    2. Fixed: Remove inital observations up to a certain number of seconds into
    the measurement (typically 30)
    3. Minima: Remove initial observations up to a local minima
    4. Optimum: Remove initial observations until an R^2^ threshold is achieved
    (success) or too few observations remain (failure).
  + **N~2~O**: Linear and quadratic models with model selection metric specified
  by user (RMSE or R^2^). Results from CO~2~ modeling can be passed to remove
  CO~2~-identified deadband from N~2~O models and/or fail N~2~O models if the
  CO~2~ model failed.
  + **Other**: Linear modeling with the option to pass results from CO~2~
  modeling to remove CO~2~-identified deadband and/or fail if the CO~2~ model
  failed.
  


