---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gasFluxR

gasFluxR provides a basic workflow to process raw data from Gasmet portable gas analyzers. The package supports data processing from import through to flux calculations. Features include:

1. Batch import and processing.
2. Identification and removal of "deadband" (period of measurement before equilibrium is reached) using CO~2~ data and user-specified parameters.
  + *In development:* Visualization of CO~2~ flux model and removed deadband.
3. Multiple-model fitting (linear and quadratic) and model selection using user-specified metric (R^2^ or RMSE).
  + Currently supported for N~2~O data only.

## Installation

You can install the development version of gasFluxR from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("redseasoils/gasFluxR")
```

## Getting Started

To model gas flux from Gasmet data using the package's functions, you'll need **a TXT file from a Gasmet portable gas analyzer**. To calculate flux values from the model, you'll also need the chamber temperature recorded during the sample, and the chamber height (ratio of chamber volume to area).

### Data Import
  
For batch import, we recommend a specific file storage structure. More on that in [File Storage](#file-storage-for-batch-processing). For now, let's import data for a single Gasmet sample (one TXT file) using [import_gasmet_data()]. 

```{r import-gasmet-file}
library(gasFluxR)
# Use an example TXT file directory
txt_file_dir <- system.file("data", "00_raw", "gas_concentration", "Site1", 
                            "20240830", package = "gasFluxR")
# And a TXT file within that directory
txt_file <- "301 BT.TXT"
# Import
gasmet_data <- import_gasmet_data(txt_file_dir, txt_file)
head(gasmet_data)
```

### Model Gas Flux

To model gas flux (ppm sec^-1^), we'll use [model_gas_flux()]. The function returns a structured list with information about the data, the model, deadband removal (if applicable), model selection (if applicable), and model "success." Model success is determined by inputs to the function, including: 

+ `min_n` : the minimum number of observations
+ `min_R2` : the minimum R^2^ value

We'll specify a minimum number of **4 observations** and a minimum R^2^ value of **0.98** for a CO~2~ model. We'll also specify a **fixed** deadband removal method to remove the first **30 seconds** of the measurement.

```{r model-co2-flux}
flux_mod_CO2 <- model_gas_flux(
  data = gasmet_data,
  ppm_var = Carbon.dioxide.CO2,
  seconds_var = seconds,
  min_n = 4,
  min_R2 = 0.98,
  deadband_opts = list(
    method = "fixed", # specify fixed deadband removal method
    fixed = list(seconds = 30) # specify more options for fixed deadband removal
  )
)
flux_mod_CO2
```

For non-CO~2~ gases, we can use the same [model_gas_flux()] function. If we want to apply the CO~2~вЂ“identified deadband to the non-CO~2~ model, we can pass the output from the CO~2~ model.

For an N~2~O model, we will also pass model selection options via the `mod_opts` argument. We will run both linear and quadratic models on the N~2~O sample, and select the final model using RMSE. For model quality control, we will specify a minimum of **4 observations** and a **minimum R^2^ of 0.1**.

```{r model-n2o-flux}
flux_mod_N2O <- model_gas_flux(
  data = gasmet_data,
  ppm_var = Nitrous.oxide.N2O,
  seconds_var = seconds,
  co2_mod = flux_mod_CO2,
  min_n = 4,
  min_R2 = 0.1,
  mod_opts = list(
    models = c("linear", "quadratic"),  # Models to fit
    selection_metric = "RMSE"  # Metric for best model selection
  )
)
flux_mod_N2O
```

For NH~3~ and CH~4~ models, we can again pass the CO~2~ model output to remove the CO~2~ identified deadband from the sample before modeling, and set our minimum number of observations and R^2^. We do not need to pass any other species-specific parameters.

```{r model-ch4-flux}
flux_mod_CH4 <- model_gas_flux(
  data = gasmet_data,
  ppm_var = Methane.CH4,
  seconds_var = seconds,
  co2_mod = flux_mod_CO2,
  min_n = 4,
  min_R2 = 0.1
)
flux_mod_CH4
```
```{r model-nh3-flux}
flux_mod_NH3 <- model_gas_flux(
  data = gasmet_data,
  ppm_var = Ammonia.NH3,
  seconds_var = seconds,
  co2_mod = flux_mod_CO2,
  min_n = 4,
  min_R2 = 0.1
)
flux_mod_NH3
```

Note that the NH~3~ flux model fails to meet our quality control parameters, so the output flags `"success"` as `FALSE` and adds a `"reason"` for model failure. Also note that `"models"` and `"metrics"` are both `NULL` as the function does not by default return models and metrics when quality control parameters are not met. To overcome this, we can tell the function to return the model regardless of quality control status via the `force` argument.

```{r model-nh3-flux-force}
flux_mod_NH3_forced <-  model_gas_flux(
  data = gasmet_data,
  ppm_var = Ammonia.NH3,
  seconds_var = seconds,
  co2_mod = flux_mod_CO2,
  min_n = 4,
  min_R2 = 0.1,
  force = TRUE
)
flux_mod_NH3_forced
```

When we set `force` to `TRUE`, we will get the model and its metrics in the output, even if the model fails quality control.

<!-- ### Calculate Gas Flux -->

<!-- 2. A spreadsheet (.xlsx) containing variable chamber dimensions and variables, including chamber temperature (В°C), collar height (cm; if applicable), chamber height (cm), and input and output tube lengths (cm). -->

<!-- ```{r import-volume} -->
<!-- # Use an example chamber variables spreadsheet -->
<!-- chamber_dir <- system.file("data", "00_raw", "chamber_volume", "Site1",  -->
<!--                            package = "gasFluxR") -->
<!-- chamber_file <- "20240830.xlsx" -->
<!-- chamber <- import_chamber_volume(chamber_dir, chamber_file) -->
<!-- head(chamber) -->
<!-- ``` -->

## File Storage for Batch Processing

Files must be structured in a standard format to batch process Gasmet data. File structure must meet the following conditions:

  1. All Gasmet TXT files must be stored in a directory in which there are no other TXT files. The files *can* be stored in subdirectories, but there must be one main directory below which the only TXT files are Gasmet files.
  2. Similarly, all spreadsheets containing chamber variables must be stored in a directory which contains no other .xlsx files (this can be the same as the TXT file directory). These files may also be stored in subdirectories. 
  
Below is an example file structure, in which "Gasmet_Data/raw" is the main directory for both Gasmet TXT files and spreadsheets:
```bash
в””в”Ђв”ЂGasmet_Data
в”‚В В  в””в”Ђв”Ђ raw
в”‚В В  в”‚В В  в”њв”Ђв”Ђ Site_A
в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 20240701
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1101.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1102.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1201.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1202.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в””в”Ђв”Ђ Chamber.xlsx
в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 20240708
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1101.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1102.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1201.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 1202.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в””в”Ђв”Ђ Chamber.xlsx
в”‚В В  в”‚В В  в””в”Ђв”Ђ Site_B
в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 20240702
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2101.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2102.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2201.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2202.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в””в”Ђв”Ђ Chamber.xlsx
в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 20240709
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2101.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2102.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2201.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в”њв”Ђв”Ђ 2202.TXT
в”‚В В  в”‚В В  в”‚В В  в”‚В В  в””в”Ђв”Ђ Chamber.xlsx
```

With the file structure above, assuming the working directory is the root directory containing the `Gasmet_Data` folder, the import functions would be:

``` {r eval=FALSE}
# Set option for TXT file directory, then import
options("gaseous.gasemt_TXT_dir" = "Data/GHG/Gasmet/raw")
import_gasmet_data()

# OR specify the path in the import function, without option setting
import_gasmet_data("Data/GHG/Gasmet/raw")

# Chamber measurements
import_chamber_volume("Data/GHG/Gasmet/raw")
```


## Flux Modeling Procedures

The function `model_gas_flux()` detects the gas species via the column name in the `ppm_var` argument (or using `gas_name` directly, if specified). The data are dispatched to modeling procedures based on species:

  + **CO~2~**: Linear model with four options for deadband (i.e. initial chamber
  disturbance) removal:
    1. None: All data are modeled
    2. Fixed: Remove inital observations up to a certain number of seconds into
    the measurement (typically 30)
    3. Minima: Remove initial observations up to a local minima
    4. Optimum: Remove initial observations until an R^2^ threshold is achieved
    (success) or too few observations remain (failure).
  + **N~2~O**: Linear and quadratic models with model selection metric specified
  by user (RMSE or R^2^). Results from CO~2~ modeling can be passed to remove
  CO~2~-identified deadband from N~2~O models and/or fail N~2~O models if the
  CO~2~ model failed.
  + **Other**: Linear modeling with the option to pass results from CO~2~
  modeling to remove CO~2~-identified deadband and/or fail if the CO~2~ model
  failed.
  


